Decentralized autonomous organization to automate governance\\ {\smaller \textbf{Final Draft - UNDER REVIEW}}

### 2016-04-05 paper/Paper.tex
second round of changes

\title{Decentralized autonomous organization to manage a trust\\ {\smaller \textbf{Final Draft - UNDER REVIEW}}}
\author{
    Christoph Jentzsch\\
    Founder \& CTO, Slock.it\\
    christoph.jentzsch@slock.it
}
\begin{document}

\begin{abstract}
This document describes the first implementation of a Decentralized Autonomous Organization (DAO) decentralizing the management of a trust. This concept, as well as the majority robbing the minority attack vector are discussed, and a solution is proposed. Furthermore, a practical implementation of this type of DAO using smart contracts written in Solidity on the Ethereum blockchain is presented and described in detail.
\end{abstract}

@@

Historically, the management of trusts has always been centralized. A group of individuals (trustees) would write a contract and combine their resources into one trust, which was then managed by a single person or company, with the trustees losing the power to directly control their money. The rise of crowdfunding (\cite{2015CFReport}), has lowered the barrier of entry to fund projects and has allowed for small contributors to participate in large projects. The money raised still carries the risk of being mismanaged by the company receiving the funding which all too often fails to follow through on its promises or outright disappears without a trace (\cite{9MostDisgracefulCrowdFundings, CrowdfundingFailTC}). Thanks to the development of Ethereum (\cite{buterin2013ethereum,Wood2014ethereum}), a blockchain technology which integrates a Turing complete programming language and the ability to process smart contracts, it is now possible to create an organization allowing the “trustees” to maintain direct control of their funds while still bundling many smaller contributions together to achieve a larger goal. We have written smart contracts (\cite{szabo1997formalizing, miller1997future}) that form a Decentralized Autonomous Organization (DAO) on the Ethereum blockchain. In this paper we detail the concept of a DAO managing a trust using a concrete example.

After explaining the concept of the DAO, we will cover the ``Majority Robbing the Minority Attack'' and propose a solution, the ``DAO split''. Finally, we will explore the smart contracts and conclude with a more detailed specification of the ``DAO split''.

@@

\section{Concept}
This section will cover the underlying concept of the DAO and what it is able to do.

After the DAO is deployed on the Ethereum blockchain, anyone is able to fund it by sending ether (the digital fuel powering the Ethereum network) to the address of the DAO’s smart contract during a period of time set aside for the initial funding. In exchange, tokens are created which represent membership as well as ownership of a portion of the DAO; these tokens are assigned to the account that sent the ether. The amount of tokens created is proportional to the amount of ether transferred. The price per token varies over time (see section \ref{TokenPrice}). The ownership of these tokens can be transferred to other accounts using transactions on the Ethereum blockchain, as soon as the funding period is over.
When deploying the contract, a minimum DAO funding goal and the period of time allowed for the initial funding are chosen. In the event the minimum funding amount is not reached at the end of the funding period, every contributor is refunded their ether. After the funding period has ended, we denote the total ether raised by $\Xi_{\text{raised}}$ and the total amount of tokens created by $T_{\text{total}}$. 
This DAO purely manages funds. In itself, it does not have the capabilities to build a product, write code or develop hardware. It requires a service provider to accomplish these and other goals, which it hires by signing off on a proposal. Every member of the DAO is allowed to submit proposals to spend a portion of the total ether raised, denoted by $\Xi_{\text{transfer}}$, on funding projects. If the proposal is approved, the ether is sent to another smart contract representing the proposed project. Such smart contracts can be parameterized and enable the DAO to interact with and influence the project being funded.
An example of such an agreement between the DAO and a project to be funded can be found in the appendix (\ref{app:SampleOffer}).

@@

This means, initially a quorum of  $20\%$ of all tokens is required for any proposal to pass. In the event $\Xi_{\text{transfer}}$ equals the amount of ether raised during the funding period plus the rewards received, then a quorum of $53.33\%$ is required.

In order to prevent proposal spam, a minimal deposit is required to be paid when creating a proposal, which gets refunded if quorum is achieved, if quorum is not achieved, the proposal deposit stays with the DAO. The value of the proposal deposit can be changed from the default value by the DAO through another proposal.

\section{Notation}
Throughout this paper, $\Xi$ always represents an amount of ether in its base unit wei. Which is defined as $1 \: \text{Wei} = 10^{-18} \: \text{Ether}$ (\cite{Wood2014ethereum}).
Similarly, DAO tokens are denoted with $T$ and always represent the amount of DAO tokens in its base unit. Which we define as $10^{-16}$ DAO token.
\section{Majority robs minority attack}
A problem every DAO has to mitigate is the ability for the majority to rob the minority. An attacker with $51\%$ of the tokens, acquired either during the funding period or purchased afterwards, could make a proposal to send all the funds to themselves. Since they would hold the majority of the tokens, they would always be able to pass their proposals.

To prevent this, the minority always needs the option to retrieve their portion of the funds. This is done by splitting the DAO into two. In the case an individual, or a group of token holders, strongly disagree with a proposal and want to retrieve their portion of the funds before the proposal gets executed, they can submit a special type of proposal to form a new DAO. The minority can then vote to move their funds to this new DAO, leaving the majority alone only able to spend their own money.
This idea originates from a blog post by Vitalik Buterin (\cite{Vitalik2015subjectivity}).

A problem this simple fix doesn't address is voter apathy: some token holders might not be actively involved in the DAO and might not follow proposals closely. An attacker could use this to their advantage. Even though the minority has the chance to retrieve their funds and split the DAO, some of them could be unaware of the situation and won't do this. For the DAO to be considered safe, it is required that inactive token holders will not lose their funds. Our proposed solution is implemented by limiting each individual DAO to a single service provider. This service provider controls the only account that can receive ether from the DAO, across all proposals. In addition, the service provider can build a “whitelist” of addresses that the DAO is allowed to send funds to. This gives the service provider of the DAO considerable power. To prevent the abuse of this power, it is possible for the DAO to vote for a new service provider, which may result in a split of the DAO as described above.

@@

The process of choosing a new service provider is as follows:
Any token holder can start a proposal to vote for a new service provider. For this proposal there is no proposal deposit to be paid, otherwise the attacker could vote for an extremely high deposit, preventing any splits. The debating period for this proposal is $10$ days.
This is $4$ days less than the minimum required for regular proposals, allowing anyone to retrieve their funds before a potentially malicious proposal goes through. There is no quorum requirement, so that every token holder has the ability to split into their own DAO. 
The debating period is used to discuss (on or off-chain) the new service provider and conduct a first vote that's non-binding. After this first vote, token holders can cast a second one to confirm its results. In the case a majority opts to  keep the original service provider, the minority can either change their vote to keep the original service provider in order to avoid a split, or inversely they can confirm their vote for a new service provider and move their portion of the funds to a new DAO.

>>>>

\title{Decentralized autonomous organization to automate governance\\ {\smaller \textbf{Final Draft - UNDER REVIEW}}}
\author{
    Christoph Jentzsch\\
    Founder \& CTO, Slock.it\\
    christoph.jentzsch@slock.it
}
\begin{document}

\begin{abstract}
This paper describes the first implementation of Decentralized Autonomous Organization (DAO) code to automate organizational governance and decision-making. The code can be used by individuals working together collaboratively outside of a traditional corporate form. It can also be used by a registered corporate entity to automate formal governance rules contained in corporate bylaws or imposed by law.  We first discuss the DAO Concept.  We then discuss minority rights, and propose a solution to a ``robbing the minority'' attack vector.  Finally, we provide a practical implementation of a first generation DAO entity using smart contracts written in Solidity on the Ethereum blockchain.
\end{abstract}

@@

Corporate entities of all kinds are governed by rules that describe permitted and proscribed conduct.  These rules may exist as private contracts (like bylaws or shareholder agreements) between corporate owners.  They may also be imposed by law in addition to or in the absence of a written agreement between participants.

Historically, corporations have only been able to act through people (or through corporate entities that were themselves ultimately controlled by people).  This presents two simple and fundamental problems.  Whatever a private contract or public law require:  (1) people do not always follow the rules and (2) people do not always agree what the rules actually require.  Collaboration without a corporate form does not solve these problems, necessarily, and it may introduce others.  In the absence of a coporate form, an explicit written agreement is substituted for unclear informal ``understandings'' and legal protections provided by a corporate form will not be available.

Rule-breaking within an organization not always obvious, and motives may not matter to stakeholders once their money has been lost.  While bad behavior may make a corporation or its management civilly or criminally liable, punishment can come as little comfort to an investor who has already lost their money.

Crowdfunding (\cite{2015CFReport}) amplifies the problem.  On the one hand, it has made it easier for small contributors to invest in large projects.  It has also made it possible for entrepeneurs to receive financial support that might not have been easily available in the past.  On the other hand, small investors remain vulnerable to financial mismanagement or outright fraud.  Because they have small stakes in a venture, they may lack power to identify problems, participate in governance decisions, or to easily recover their investment.  (\cite{9MostDisgracefulCrowdFundings, CrowdfundingFailTC}).  At the same time, corporate leadership and managment may be accused of malfeasance or mismanagement when they believe that they have acted in good faith, based on their understanding of their obligations and interpretation of applicable rules.

We present a potential solution using Ethereum, (\cite{buterin2013ethereum,Wood2014ethereum}) a blockchain technology which integrates a Turing complete programming language with smart contract processing functionality.  We illustrate a method that for the first time allows the creation of organizations in which (1) ?participants?  maintain direct real-time control of contributed funds and (2) where governance rules are formalized and automated using software.  Specifically, we have written generic smart contract code (\cite{szabo1997formalizing, miller1997future}) that can be used to form a Decentralized Autonomous Organization (DAO) on the Ethereum blockchain. In this paper we explain how the DAO code works.  We do so by focusing on some basic formation and governance features, including formation, funding and voting rights.

We first describe the DAO's formation and basic functionality.  We then discuss minority owner rights and propose a solution to the ``Majority Robbing the Minority Attack'' problem, the ``DAO split''. We then explore the smart contract code in detail, and conclude with an explanation and detailed specification of the ``DAO split''.

@@

A word of caution, at the outset:  the legal status of DAOs remains the subject of active and vigorous debate and discussion.  Not everyone shares the same definition.  Some have said that they are autonomous code and can operate independently of legal systems; others have said that they must be owned or operate by humans or human created entities. There will be many uses cases, and the DAO code will develop over time.  Ultimately, how a DAO functions and its legal status will depend on many factors, including how DAO code is used, where it is used, and who uses it.  We do not speculate about the legal status of DAOs worldwide.  This paper is not intended to offer legal advice or conclusions.  Anyone who uses DAO code will do so at their own risk.

\section{Dao Concept | Dao Practice}

DAO code is written in the ``Solidity'' programming language.  A DAO is activated by deployment on the Ethereum blockchain.

Once deployed, the DAO code requires ``ether'' to engage in transactions on Ethereum.  (Ether is the digital fuel that powers the Ethereum network).  Without ether, a DAO can not do very much and a DAO's first order of business is to receive ether.  After the DAO code is deployed, ether may be sent to the DAO?s smart contract address during an initial funding period which is defined in the DAO code.

In exchange for ether, the DAO code creates tokens that are then assigned to the account of the person who sent the ether.  A token grants its holder voting and ownership rights.  The number of tokens created is proportional to the amount of ether transferred.  Token price varies over time (see section \ref{TokenPrice}). Token ownership is freely transferable on the Ethereum blockchain, when the funding period is over.

A minimum DAO funding goal and funding time-period are set as parameters in the DAO code prior to deployment.  If the minimum funding amount is not reached at the close of the funding period, all ether is returned. After the funding period has ended, we denote the total ether raised by $\Xi_{\text{raised}}$ and the total amount of tokens created by $T_{\text{total}}$.

The DAO stores ether and transmits it based on the DAO code. It does not do much else.  It cannot build a product, write code or develop hardware.  It requires what we have called a "Service Provider'' to accomplish these and other goals.  The DAO engages a Service Provider by accepting a Service Provider proposal.

DAO token holders may also submit proposals to use the DAO's ether, denoted by $\Xi_{\text{transfer}}$, for a variety of purposes.  If a proposal is approved, the DAO tranmits ether to a smart contract representing the proposed project. Such smart contracts can be parameterized and enable the DAO to interact with and influence the project being funded.
An example of such an agreement between the DAO and a project to be funded can be found in the appendix (\ref{app:SampleOffer}).

@@

This means, initially, a quorum of  $20\%$ of all tokens is required for any proposal to pass. In the event $\Xi_{\text{transfer}}$ equals the amount of ether raised during the funding period plus the rewards received, then a quorum of $53.33\%$ is required.

In order to prevent ``proposal spam'', a minimal deposit is required to be paid when creating a proposal, which gets refunded if quorum is achieved. If quorum is not achieved, the DAO keeps the proposal deposit. The value of the proposal deposit can be changed from the default value by the DAO through another proposal.

\section{Notation}
Throughout this paper, $\Xi$ always represents an amount of ether in its base unit wei. This is defined as $1 \: \text{Wei} = 10^{-18} \: \text{Ether}$ (\cite{Wood2014ethereum}).
Similarly, DAO tokens are denoted with $T$ and always represent the amount of DAO tokens in its base unit. We define this as $10^{-16}$ DAO token.
\section{Majority robs minority attack}
Minority owner rights can be a problem in any corporate form.  Minority rights may be protected or addressed by provisions in corporate governance documents or by statute or judge-made law.  But many of these solutions fail because minority owners may lack voting rights or the ability to ``vote with their feet" and easily retrieve their capital.  We present a solution to this problem in the DAO code.

A problem every that DAO has to mitigate is the ability for the majority to rob the minority by changing governance and ownership rules after DAO formation.  For example, an attacker with $51\%$ of the tokens, acquired either during the funding period or purchased afterwards, could make a proposal to send all the funds to themselves. Since they would hold the majority of the tokens, they would always be able to pass their proposals.

To prevent this, the minority must always have the ability to retrieve their portion of the funds.  Our solution is to allow the DAO to split into two. If an individual, or a group of token holders, disagree with a proposal and want to retrieve their portion of the funds before the proposal gets executed, they can submit a special type of proposal to form a new DAO. The minority can then vote to move their funds to this new DAO, leaving the majority alone only able to spend their own money.
This idea originates from a blog post by Vitalik Buterin (\cite{Vitalik2015subjectivity}).

A problem this simple fix doesn't address is voter apathy: some token holders might not be actively involved in the DAO and might not follow proposals closely. An attacker could use this to their advantage. Even though the minority has the chance to retrieve their funds and split the DAO, some could be unaware of the situation and fail to act. For the DAO to be considered safe, it is required that inactive token holders must also be protected from losing their tokens. Our proposed solution is implemented by limiting each individual DAO to a single service provider. This service provider controls the only account that can receive ether from the DAO, across all proposals. In addition, the service provider can build a “whitelist” of addresses that the DAO is allowed to send funds to. This gives the service provider of the DAO considerable power. To prevent the abuse of this power, it is possible for the DAO to vote for a new service provider, which may result in a split of the DAO as described above.

@@

The process of choosing a new service provider is as follows:
Any token holder can submit a proposal for a new service provider. In this case, no proposal deposit is required, because an attacker could vote for an extremely high deposit, preventing any splits. The debating period for this proposal is $10$ days.
This is $4$ days less than the minimum required for regular proposals, allowing anyone to retrieve their funds before a potentially malicious proposal goes through. There is no quorum requirement, so that every token holder has the ability to split into their own DAO. 
The debating period is used to discuss (on or off-chain) the new service provider and conduct a first vote that's non-binding. After this first vote, token holders can cast a second one to confirm its results. In the case a majority opts to  keep the original service provider, the minority can either change their vote to keep the original service provider in order to avoid a split, or inversely they can confirm their vote for a new service provider and move their portion of the funds to a new DAO.
